### 2020-05-31 writeup/experiments.tex
wip
sec 4

@@

% \begin{table}
% \begin{center}
%   \begin{tabular}{c c c c c}
%   \toprule
%     \multirow{2}{*}{{\bf Threshold}}  & \multicolumn{2}{c}{{\bf Text classification}} & \multicolumn{2}{c}{{\bf ChestX-ray14}} \\
%     & Precision &  Recall & Precision &  Recall \\
%     \cmidrule(lr){1-1} \cmidrule(lr){2-3} \cmidrule(lr){4-5}
%     0.0 & 0.596 & 1.000 & 0.593 & 1.000 \\
%     0.1 & 0.756 & 0.388 & 0.738 & 0.462 \\
%     0.2 & 0.919 & 0.065 & 0.875 & 0.044 \\	
%     0.3 & 1.000 & 0.004 &     - &     - \\
%   \bottomrule
%   \end{tabular}
% \end{center}
% \caption{Ablation study on when should use MTL via different source/target task accuracy.}
% \label{tab:mtl_better_than_stl}
% \end{table}

\textbf{Data size.}
Recall that 
We measure the data efficiency ratio on the sentiment analysis tasks.
We find that by performing multi-task learning, only $39\%$ of the labeled data is needed to achieve comparable performance to single-task learning over all six tasks.
If TREC is not included, we see that only $25\%$ of the labeled data is needed.

\textbf{Covariate shift.}
We validate that the performance of aligning task covariances depends on the size of source task data.
In Figure \ref{fig_covariate}, we observe that the benefit from aligning task covariances becomes more significant as we increase the size of the source task dataset.

\begin{figure}[!ht]
	\centering
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.975\textwidth]{figures/ratio_alignment_norm_diff_all.pdf}
		\caption{Averaged}
	\end{subfigure}
	\caption{The performance of aligning task covariances depends on data size. As the ratio between source task data size and target task data size increases, the performance improvement from aligning task covariances increases.}
	\label{fig_covariate}
\end{figure}

%\subsection{Validating Theoretical Implications}

%\textbf{Contributing causes of negative transfer.}

%\textbf{Implications for positive learning.}

%\textit{Transferring from a high accuracy to low accuracy.}%%%%%%%%

% \begin{table}
% 	\begin{center}
% 		\begin{tabular}{c c c c c}
% 		\toprule
% 		\multirow{2}{*}{{\bf Models}} & \multicolumn{2}{c}{\begin{minipage}{1.1in}\begin{center}
% 		MR, SST, SUBJ, CR, MPQA, TREC\end{center}\end{minipage}} & \multicolumn{2}{c}{\begin{minipage}{1.1in}\begin{center}MR, SST, SUBJ, CR, MPQA\end{center}\end{minipage}} \\
% 		\cmidrule(lr){2-3} \cmidrule(lr){4-5}
% 		& {\bf Standard} & {\bf Alignment} & {\bf Standard} & {\bf Alignment} \\
% 		\midrule
% 		{\bf MLP}  & > 100\% & 39\% & 25\% & 25\% \\
% 		{\bf LSTM} & 36\% & 36\% & 28\% & 25\% \\
% 		% {\bf CNN}  & 76\% & - & 32\% & -\\
% 		\bottomrule
% 		\end{tabular}
% 	\end{center}
% 	\caption{Taskonomy experiment.}
% 	\label{tab:taskonomy}
% 	\end{table}

%\begin{figure}
%	\centering
%	\includegraphics[width=0.5\textwidth]{figures/ratio_alignment_mlp.pdf}
%	\caption{Covariate shift experiment.}
%\end{figure}

>>>>

\textbf{MTL improves labeled data efficiency.}
We measure the data efficiency ratio on the sentiment analysis tasks.
We find that by performing multi-task learning, only $39\%$ of the labeled data is needed to achieve comparable performance to single-task learning over all six tasks.
If TREC is not included, we see that only $25\%$ of the labeled data is needed.

@@

\textbf{Data size.}
Recall that Proposition \ref{prop_data_size} shows that increasing the data size of the source task does not always improve the performance of MTL for the target task.
In Figure \ref{fig_ab_data}, we observe that for the task pair MR and SST, increasing the data size of MR is helpful initially, but then gets worse.

\textbf{Covariate shift.}
We validate that the performance of aligning task covariances depends on the size of source task data.
In Figure \ref{fig_ab_cov}, we observe that the benefit from aligning task covariances becomes more significant as we increase the size of the source task dataset.

\begin{figure}[!t]
	\centering
		\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.975\textwidth]{figures/ratio_norm_sst_trec_lstm.pdf}
		\caption{Data size}
		\label{fig_ab_data}
	\end{subfigure}
	\begin{subfigure}[b]{0.33\textwidth}
		\centering
		\includegraphics[width=0.975\textwidth]{figures/ratio_alignment_norm_diff_all.pdf}
		\caption{Covariate shift}
		\label{fig_ab_cov}
	\end{subfigure}
	\caption{The performance of aligning task covariances depends on data size. As the ratio between source task data size and target task data size increases, the performance improvement from aligning task covariances increases.}
	\label{fig_ablation}
\end{figure}

### 2020-05-31 writeup/proof_main.tex

Given the optimal ratio $\hat v$, let $\hat{M} = \hat{v} \Sigma_1^{1/2}\Sigma_2^{-1/2}$ denote the weighted covariate shift matrix. Denote by ${\hat\lambda}_1\ge {\hat\lambda}_2 \ge \dots \ge {\hat\lambda}_p$ the eigenvalues of $\hat{M}^{\top}\hat{M}$. Let $(\hat a_1, \hat a_2)$ be the solutions to the following deterministic equations
	\be
		 \hat a_1 +  \hat a_2 = 1- \frac{1}{\rho_1 + \rho_2},\quad  \hat a_1 + \frac1{\rho_1 + \rho_2}\cdot \frac1p\sum_{i=1}^p \frac{ \hat \lambda_i^2 \hat a_1}{ \hat \lambda_i^2\hat a_1 +  \hat a_2} = \frac{\rho_1}{\rho_1 + \rho_2}.\label{eq_a2} \\
		 \ee
		 After obtaining $(\hat a_1,\hat a_2)$, we can solve the following linear equations to get $(\hat a_3,\hat a_4)$:
\begin{gather}
		\left(\rho_2 \hat a_2^{-2}- \hat b_0\right)\cdot \hat  a_3 - \hat b_1 \cdot \hat a_4
		=\hat b_0, \quad \left(\rho_1\hat a_1^{-2} - \hat b_2  \right)\cdot \hat a_4 - \hat b_1 \cdot \hat a_3 =\hat b_1 .\label{eq_a3} 
%		\left(\frac{n_1}{\hat a_1^2} -  \sum_{i=1}^p \frac{\hat \lambda_i^4   }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_4 -\left(\sum_{i=1}^p \frac{\hat \lambda_i^2  }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_3
%		= \sum_{i=1}^p \frac{\hat \lambda_i^2 }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }. \label{eq_a4}
	\end{gather}
where we denoted
$$\hat b_k:= \frac1{p}\sum_{i=1}^p \frac{\hat \lambda_i^{2k}}{ (\hat a_2 +\hat  \lambda_i^2\hat a_1)^2  },\quad k=0,1,2.$$
Then we introduce the following matrix
\be\label{defnpihat}\Pi = \frac{\rho_1^2}{(\rho_1 + \rho_2)^2}\cdot{\hat M} \frac{(1 + \hat a_3)\id +\hat  a_4 \hat {M}^{\top}\hat {M}}{(\hat a_2 +\hat  a_1 \hat {M}^{\top}\hat {M})^2} \hat {M}^{\top},\ee
which is defined in a way such that {\it in certain sense} it is the asymptotic limit of the random matrix 
$$\hat v\Sigma_1^{1/2} (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} \Sigma_2 (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_1^{1/2}.$$ 
We introduce two factors that will appear often in our statements and discussions:
$$\al_-(\rho_1):=(1- \rho_1^{-1/2})^2,\quad \al_+(\rho_1):=(1 + \rho_1^{-1/2})^2.$$ 
In fact, $\al_-^2$ and $\al_+^2$ correspond to the largest and smallest singular values of $Z_1/\sqrt{n_1}$, respectively, as given by the famous Mar{\v c}enko-Pastur law \cite{MP}. In particular, as $\rho_1$ increases, both $\al_-$ and $\al_+$ will converge to 1 and $Z_1/\sqrt{n_1}$ will be more close to an isometry. Finally, we introduce the error term  
\be\label{eq_deltaextra}\delta:=\frac{\al_+^2(\rho_1) - 1 }{\al_-^{2}(\rho_1)\lambda_{\min}^2(\hat M)} \cdot  \norm{\Sigma_1^{1/2}(\beta_s - \hat{w}\beta_t)}^2,\ee
where $\lambda_{\min}(\hat M)$ is the smallest singular value of $\hat M$. Note that this factor converges to 0 as $\rho_1\to \infty$.

>>>>

Given the optimal ratio $\hat v$, let $\hat{M} = \hat{v} \Sigma_1^{1/2}\Sigma_2^{-1/2}$ denote the weighted covariate shift matrix. Denote by ${\hat\lambda}_1\ge {\hat\lambda}_2 \ge \dots \ge {\hat\lambda}_p$ the eigenvalues of $\hat{M}^{\top}\hat{M}$. Let $(\hat a_1, \hat a_2)$ be the solutions to the following deterministic equations
	\be
		 \hat a_1 +  \hat a_2 = 1- \frac{1}{\rho_1 + \rho_2},\quad  \hat a_1 + \frac1{\rho_1 + \rho_2}\cdot \frac1p\sum_{i=1}^p \frac{ \hat \lambda_i^2 \hat a_1}{ \hat \lambda_i^2\hat a_1 +  \hat a_2} = \frac{\rho_1}{\rho_1 + \rho_2}.\label{eq_a2} \\
		 \ee
		 After obtaining $(\hat a_1,\hat a_2)$, we can solve the following linear equations to get $(\hat a_3,\hat a_4)$:
\begin{gather}
		\left(\rho_2 \hat a_2^{-2}- \hat b_0\right)\cdot \hat  a_3 - \hat b_1 \cdot \hat a_4
		=\hat b_0, \quad \left(\rho_1\hat a_1^{-2} - \hat b_2  \right)\cdot \hat a_4 - \hat b_1 \cdot \hat a_3 =\hat b_1 .\label{eq_a3} 
%		\left(\frac{n_1}{\hat a_1^2} -  \sum_{i=1}^p \frac{\hat \lambda_i^4   }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_4 -\left(\sum_{i=1}^p \frac{\hat \lambda_i^2  }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_3
%		= \sum_{i=1}^p \frac{\hat \lambda_i^2 }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }. \label{eq_a4}
	\end{gather}
where we denoted
$$\hat b_k:= \frac1{p}\sum_{i=1}^p \frac{\hat \lambda_i^{2k}}{ (\hat a_2 +\hat  \lambda_i^2\hat a_1)^2  },\quad k=0,1,2.$$
Then we introduce the following matrix
\be\label{defnpihat}\Pi = \frac{\rho_1^2}{(\rho_1 + \rho_2)^2}\cdot{\hat M} \frac{(1 + \hat a_3)\id +\hat  a_4 \hat {M}^{\top}\hat {M}}{(\hat a_2 +\hat  a_1 \hat {M}^{\top}\hat {M})^2} \hat {M}^{\top},\ee
which is defined in a way such that {\it in certain sense} it is the asymptotic limit of the random matrix 
$$\hat v\Sigma_1^{1/2} (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} \Sigma_2 (\hat{v}^2 X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_1^{1/2}.$$ 
We introduce two factors that will appear often in our statements and discussions:
$$\al_-(\rho_1):=(1- \rho_1^{-1/2})^2,\quad \al_+(\rho_1):=(1 + \rho_1^{-1/2})^2.$$ 
In fact, $\al_-^2$ and $\al_+^2$ correspond to the largest and smallest singular values of $Z_1/\sqrt{n_1}$, respectively, as given by the famous Mar{\v c}enko-Pastur law \cite{MP}. In particular, as $\rho_1$ increases, both $\al_-$ and $\al_+$ will converge to 1 and $Z_1/\sqrt{n_1}$ will be more close to an isometry. Finally, we introduce the error term  
\be\label{eq_deltaextra}\delta\equiv \delta(\hat v):=\frac{\al_+^2(\rho_1) - 1 }{\al_-^{2}(\rho_1)\lambda_{\min}^2(\hat M)} \cdot  \norm{\Sigma_1^{1/2}(\beta_s - \hat{v}\beta_t)}^2,\ee
where $\lambda_{\min}(\hat M)$ is the smallest singular value of $\hat M$. Note that this factor converges to 0 as $\rho_1\to \infty$.

### 2020-05-31 writeup/transfer.tex

We measure model dissimilarity as $\norm{\beta_1 - \beta_2}^2$, which is the distance between source and target in the isotropic model.
We derive a sharp threshold when positive transfer transitions to negative transfer, as model dissimilarity increases.
%Based on Theorem \ref{thm_model_shift}, we derive the transition threshold in the following proposition.
Define the following functions
\begin{align*}
	\Psi(\beta_1, \beta_2) = {\ex{\bignorm{\beta_1 - \beta_2}^2}} / {\sigma^2},  \Phi(\rho_1, \rho_2) = \frac{(\rho_1 + \rho_2 - 1)^2}{\rho_1 (\rho_1 + \rho_2) (\rho_2 - 1)}.
\end{align*}
Let $\gamma_{-} = (1+o(1))\cdot(1 - \rho_1^{-1/2})^4$ and $\gamma_{+} = (1+o(1))(1 + \rho_1^{-1/2})^4$.
Concretely, if $\rho_1 > 40$, then $\gamma_{-} < 2$ and $\gamma_{+} > 1/2$.

@@

\begin{proposition}[Unnecessary source data]\label{prop_data_size}
	In the isotropic model, assume that $\rho_1 > 40$ and $\rho_2 > 500$ are fixed constants, $\Psi(\beta_1, \beta_2) > 2/(\rho_2 - 1)$.
	We have that
	\squishlist
		\item If $\rho_1 < \frac{\rho_2-2}{\gamma_+ \cdot \Psi(\beta_1, \beta_2) (\rho_2 - 1) - 1}$, then w.h.p. $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\rho_1 > \frac{\rho_2-2}{\gamma_- \cdot \Psi(\beta_1, \beta_2) (\rho_2 - 3) - 1}$, then w.h.p. $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\squishend
\end{proposition}
Proposition \ref{prop_data_size} shows that as the source data size increases, we observe a transition from positive to negative transfer.
The proof of Proposition \ref{prop_data_size} is similar to Proposition \ref{prop_dist_transition}, and can be found in Appendix \ref{app_proof_31}.

>>>>

We measure model dissimilarity as $\norm{\beta_1 - \beta_2}^2$, which is the distance between source and target in the isotropic model.
We derive a sharp threshold when positive transfer transitions to negative transfer, as model dissimilarity increases.
%Based on Theorem \ref{thm_model_shift}, we derive the transition threshold in the following proposition.
Define the following functions
\begin{align*}
	\Psi(\beta_1, \beta_2) = {\ex{\bignorm{\beta_1 - \beta_2}^2}} / {\sigma^2},  \quad \Phi(\rho_1, \rho_2) = \frac{(\rho_1 + \rho_2 - 1)^2}{\rho_1 (\rho_1 + \rho_2) (\rho_2 - 1)}.
\end{align*}
Let $\gamma_{-} = (1-\oo(1))\cdot(1 - \rho_1^{-1/2})^4$ and $\gamma_{+} = (1+\oo(1))\cdot(1 + \rho_1^{-1/2})^4$.
Concretely, if $\rho_1 > 40$, then $\gamma_{-} < 2$ and $\gamma_{+} > 1/2$.

@@

\begin{proposition}[Unnecessary source data]\label{prop_data_size}
	In the isotropic model, assume that $\rho_1 > 40$ and $\rho_2 > 150$ are fixed constants, $\Psi(\beta_1, \beta_2) > 2/(\rho_2 - 1)$.
	We have that
	\squishlist
		\item If $\rho_1 < \frac{\rho_2-2}{\gamma_+ \cdot \Psi(\beta_1, \beta_2) (\rho_2 - 1) - 1}$, then w.h.p. $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\rho_1 > \frac{\rho_2-2}{\gamma_- \cdot \Psi(\beta_1, \beta_2) (\rho_2 - 3) - 1}$, then w.h.p. $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\squishend
\end{proposition}
Proposition \ref{prop_data_size} shows that as the source data size increases, we observe a transition from positive to negative transfer.
The proof of Proposition \ref{prop_data_size} is similar to Proposition \ref{prop_dist_transition}, and can be found in Appendix \ref{app_proof_31}.
